Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-07 doc/thesis/report.tex
Corrected entirely. Figures will be changes because they are not appropriate

\documentclass[oneside,12pt,a4paper,final]{book}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\chapter{Abstract}
\paragraph{}
The information technology is advancing exponentially and is having bigger impact on our lives as time advances. The next revolution in computing is called the \gls{iot}. It is basically the internet we know nowadays except that physical objects (fridge, television, air conditioning...) will be the major players in this complex system. In this context, the project builds on this idea a test bed for smart home energy optimization. The project allows users to monitor their energy consumption via their smart phone and also the users can control their appliances by turning on and off appliances remotely.
\paragraph{}
The system is divided into three main components. Each of them is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} and other technologies that will be discussed in the thesis report.
\paragraph{}
At this stage, the system allows monitoring, but turning on and off appliances is not yet there mainly because of the absence of the equipment but the communication for that purpose is there.

\chapter{Résumé}
\paragraph{}
La technologie de l'information progresse de façon exponentielle et son impact sur nos vies prend de l'importance plus le temps avance. La prochaine révolution de l'informatique est appelé \glsreset{iot} \gls{iot}. Il s'agit essentiellement de Internet tel que nous la connaissons aujourd'hui, sauf que les objets physiques  seront les principaux acteurs de ce système. Ainsi, le projet s'appuie sur cette idée et présente un prototype qui a pour but l'optimisation intelligente de l'énergie dans les domiciles. Le projet permet aux utilisateurs de suivre de près leur consommation d'énergie via leur smartphone et permet de contrôler leurs appareils en allumant et éteignant les appareils à distance.
\paragraph {}
Le système est divisé en trois composantes principales. Chacun d'entre eux  utilise différentes technologies telles que \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} et d'autres technologies qui seront présentées dans le rapport de thèse.
\paragraph{}
A ce stade, le système permet de surveiller, mais controler les appareils n'est pas encore mise en place principalement en raison de l'absence de matériel, mais l'infrastructure de communication pour ce but là est en place.
\setcode{utf8}

@@

\section{Importance of Study}
\paragraph{}
The global electrical grid one the bases that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man. \cite{ref1}. It a widespread machine, distributed but interconnected system that provides power to households, companies and other organizations and therefore the world's economy is counting on it. The electrical grid knows numerous problems such as blackouts, current instabilities that costs billions of dollars yearly to both the utilities, utilities and insurances. The grid suffers from the problem of peak hours where energy consumption where in many case the one single utility cannot provide such energy.
\paragraph{}
The \gls{sg} comes to provide solutions to the problems that suffers from the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and is and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arises in the traditional grid such as the energy loss in the energy transmission lines and also in the overload of the electrical lines.
\paragraph{}
Another revolution is about to begin and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously without the guidance of humans to performs their tasks. Cisco predicts that the number of devices to be connected in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion. That make 6.58 devices connected for each person in the world whereas the number of devices connected in 2010 12.5 billion. This clearly shows that the shift to the \gls{iot} has already started.
\paragraph{}
\glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}. The integration of the \gls{sg} and and \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The smart grid as a set of philosophical ideas counts on a set of novel functionalities. Once of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa. This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information. Numerous ideas are based on this two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and other technologies which make the Internet of things a very heterogeneous and complex system \cite{ref4}. One of the challenges in \gls{iot} is to deals with such heterogeneity and take the most out of it in order to make the communication as smooth and robust as possible. This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication and protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of the \gls{dr}. From the consumer's side or more precisely the household's side, there are smart home energy management software in the market nowadays but none of these products offers integration with \gls{sg}. To have such an integration with \gls{sg}, there is a need to first build this two-way communication and make it possible for any appliance to report its real-time energy consumption and also to be able to be controlled remotely with this two-way communication. Such infrastructure is missing for the time being and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
\paragraph{}
As \gls{sg} counts on numerous technologies to achieve its main goals. This usage of different technologies imposes another problem which is the heterogeneity of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. This means that there is a need to build a software within the whole system that would constitute an important part of it and will take care of making the communication and processes technology-transparent and give the possibility to all components of the system to work hand-in-hand in order their common goals independently of the data link layer technology (Ethernet, Wifi or Zigbee) they are counting on, the network layer technology (\gls{ipv4} or \gls{ipv6}).

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and also solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of things by making it communicate with any device in the Internet and having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational and energy rich.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on this possibility. To do it, there was a need to first introduce \gls{6lowpan} within every device in the \gls{wsn}. This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}. This communication will go in the two directions and based on that the two-way communication will have to be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an additional component that will track all communication packets and change them to fit the needs of the destination component. This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
The thesis aims to solves the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate one cannot solve without having to solve the other. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other network only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem. This is why the two-way communication was created and then the heterogeneity problem had to be solved in order to see everything working.
The system is divided into four components:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The \gls{tinyos} program for sensor transmission and appliance request handler
\item The heterogeneity handler and middleman gateway server
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four component exist to solve part of the problem. Each of these components will be discussed extensively in the Chapter 4 from the architectural aspect of the component and in Chapter 5 from the implementation level of the system. 
\paragraph{}
It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation point in Chapter 5. The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment was omitted to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show concepts of the \glsreset{iot}\gls{iot} which are that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically. The consumption of data by is shown by external nodes in the \gls{iot} sending request to the sensor node in the \gls{wsn} in order to control and appliance's status i.e. turn on the television. This will prove the consumption of information by the nodes in the \gls{iot}.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful, robust and reliable and that the heterogeneity of the system is made transparent to the nodes in the system by the creation of the software tool residing in the gateway which is between the \gls{iot} and the outside world. This software tool is seen as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will able to show the two-way communication in action and will also show that the heterogeneity does not form a barrier by its successful handling.

\section{Outline of the Thesis}
The thesis is is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
\paragraph{Chapter 1}
It is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic. This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 
\paragraph{Chapter 2}
It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}. It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
\paragraph{Chapter 3}
It is dedicated to the \glsreset{iot}\gls{iot}. It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.
\paragraph{Chapter 4}
It will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
\paragraph{Chapter 5}
This chapter will mostly discuss the implementation system. It will start by presenting the hardware used in this project, then move to the technology enablers. Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
\paragraph{Chapter 6}
It will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system. The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
\paragraph{Chapter 7, 8}
These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.

@@

The actual electrical grid is based on a technology that is at least one century old \cite{ref5}. As a result, there is a need to change as to meet to 21st century requirement such as having more green energy in order to reduce the pollution caused mainly by the use of fossil energy. As a result, research has introduced a new technology called \glsreset{sg}\gls{sg}. Although it is still a research project and numerous issues related to \gls{sg} are left unanswered, nations across the globe started to deploy \gls{sg} project. It is also noted that in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.

@@

The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.  This distribution by itself enables the power grid this to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer but households can now produce and even sell the extra power. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producer and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide power to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}

@@

The \glsreset{iot}\gls{iot} has known high acceptance by the players in the Internet. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to read 50 billion connected devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and these benefits will pull \gls{sg} to be part of it. Consequently, research has taken this direction which is to integrate the \gls{sg} within the \gls{iot} as most aspects of \gls{sg} will be enhanced by \gls{iot} \cite{ref8}.

@@

\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power lines, WI-FI, Zigbee and other technologies \cite{ref8}. The smart grid's architecture based on \gls{iot} will be divided into three layers:

@@

\item Perception layer: Devices will autonomously perform their tasks such as sensing and executing orders.
\item Communication layer: The communication infrastructure is defined here and issues related to how to communicate information is handled.
\item Application layer: Applications can be built to make use of the underlying data.

@@

Some of the research is discussing the need to replace Zigbee by Wi-Fi as the Wi-Fi offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10} but still Zigbee has the sole advantage of very low energy consumption which is something required in \gls{sg} and \gls{iot}, because in systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices, one cannot manage to move for the sole purpose of providing energy often. 

@@

From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues. An \gls{iot} gateway system that is based on Zigbee and GPRS protocols \cite{ref11}. This enables then to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network. Another solution to the heterogeneity problem is proposed where the solution is to present a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices \cite{ref12}. This protocol is platform-independent, low energy consumer. Some researchers claim that the heterogeneity is mainly coming from the fact that there are different \gls{wsn} devices that do not necessarily use the same standards and protocols. They propose to move all the \glspl{wsn} to all-IP networks as this would remove most of the heterogeneity but there are already deployed legacy \glspl{wsn} \cite{ref13}. They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}
\section{The \gls{sg} Two-Way Communication}

\chapter{Internet Of Things}

\section{The actual Internet}
\section{Introduction to \gls{iot}}
\section{Importance of \gls{iot}}
\section{The Impact of \gls{iot}}
\section{The Benefits of \gls{iot}}
\section{The Future of Internet}

\chapter{System Architecture}
\paragraph{}
The thesis is based on the study and creation of a test-bed system whose goal is to integrate smart home with the \glsreset{sg}\gls{sg} and all within the \glsreset{iot}\gls{iot}. To do so, we first define the system's objectives and goals which do align  with the thesis's problem statement and purpose of study. Then each part of the system is presented, and a section of this chapter is dedicated to each part explaining the architecture of each component and how it aligns and integrates with the whole system.
\section{System Objectives}
\paragraph{Objective 1}
The goal of the system is to present a working test-bed that is able to interact with any appliance in the smart home. The system makes this possible by making all appliances part of the \glsreset{wsn}\gls{wsn} where each sensor node is addressable and may be used to control the appliance.
\paragraph{Objective 2}
Another system objective is to be able to monitor in real-time the consumption of each appliance and send this to a central repository where this data is to be stored. This will enable the sensor node to keep track of the appliance it is controlling and monitoring and collect consumption data that is sent and stored in a \gls{rdbms}.
\paragraph{Objective 3}
The third objective is to make this system hide its heterogeneity from the outside world and also from different component belonging to it. This means that the system should be able to make any two components communicate independently of the technology they are using and also making each component think the other is using the same technology as his.

\section{System Components}
\paragraph{}
The system is divided into nine components where every component is in charge of a part of the system's objectives. When these nine components work hand-in-hand, they achieve all the system's and project's objectives and goals. The components that are defined for this system are the following:

\begin{itemize}
\item Full \gls{ipv6} \glsreset{wsn}\gls{wsn}
\item Mote sensor data collection TCP client
\item Mote appliance control TCP server
\item Gateway packet transformation process
\item \gls{ipv6} over USB tunnel
\item mote \gls{ipv6} backstation
\item Sensor data storage process
\item HTTP sensor data sender
\item Mobile client
\end{itemize}

>>>>

\documentclass[oneside,12pt,a4paper,final]{book}
%\usepackage[margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
\usepackage[titletoc]{appendix}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\chapter{Abstract}
\paragraph{}
Information Technology is advancing exponentially and largely impacting our daily lives. The current evolution in computing is the \gls{iot}. It is smoothly evolving from an Internet of people to and Internet of Things. By 2020, it is expected to have 50 billion "Things" connected to the Internet. 
\paragraph{}
However such a migration induces a strong level of complexity when handling interoperability between heterogeneous Things, i.g. Wireless Sensors, RFID, and mobile devices.
\paragraph{}
In this thesis, we focus on the  integration of wireless sensors into \gls{iot}. This is q prologue for the future integration of \gls{sg} into \gls{iot}. In this work, we deployed a smart building testbed for energy optimization, and for integration into \gls{sg}\glsreset{sg}.
\paragraph{}
The testbed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
\paragraph{}
The system is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se} and  \gls{http}.

\chapter{Résumé}
\paragraph{}
La technologie de l'information progresse de façon exponentielle et prend largement plus impact sur nos vies quotidiennes. L'évolution actuelle de l'informatique est appelé \glsreset{iot}\gls{iot}. il est prévu d'avoir en 2020 50 milliards "choses" connecté à l'Internet.
\paragraph{}
Cependant, une telle migration induit un fort niveau de complexité lorsque l'interopérabilité entre en jeu entre differentes choses hétérogènes, à savoir capteurs sans fil, RFID et l'informatique mobile.
\paragraph{}
Dans cette thèse, nous nous concentrons sur l'intégration de capteurs sans fil dans \gls{iot}. Cela conduit vers une future intégration de Smart Grid (SG) dans \gls{iot}. Au cours de ce projet, nous avons déployé un banc d'essai intelligente du bâtiment pour l'optimisation de l'énergie.
\paragraph{}
Ce système d'information utilise différentes technologies  telles que TinyOS (TinyOS), \glsreset{nesc}\gls{nesc}, Java Standard Edition (J2SE), HTTP.

\setcode{utf8}

@@

\section{Importance of Study}
\paragraph{}
The global electrical grid is the basis that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}. It is a widespread and distributed machine that provides power to households, companies and organizations. Therefore the whole world's economy depends on it. The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances. The grid suffers from the problem of peak hours where energy consumption is at its highest level and that one single utility cannot provide.
\paragraph{}
The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources, reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
\paragraph{}
Another revolution began, and it is taking place in the Information Technology. It is called the \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously without the guidance of humans, to perform their tasks. Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003. That would make 6.58 devices connected for each person in the world whereas that same ratio was 0.07 17 years earlier \cite{ref3}. This clearly shows that the shift to the \gls{iot} will be exponential.
\paragraph{}
\glsreset{sg}\gls{sg} should benefit from the \gls{iot}. The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}. This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy use. This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e. the Internet.

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The Smart Grid needs a set of novel functionalities. One of these ideas is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa. This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information. Numerous ideas assume the existence of the two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}. One of the challenges in \gls{iot} is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved. This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's/household's perspective or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}. To have such an integration with \gls{sg}, there is a need for the two-way communication, in order to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. This use of different technologies comes with a problem which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system whose main goal is to take care of making the communication technology-transparent. This component will give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid. In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it. To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}. This would make every sensor node (mote) identifiable in the network, ready to communicate and be part of the \gls{iot}. This communication will go in both directions and based on \gls{6lowpan}; the two-way communication will be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the targeted component. This will make of this additional component a middleman and that would make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
This study aims at solving the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other. Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system. This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The heterogeneity handler and middleman gateway server
\item The middle-ware
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four components is there to solve part of the problem. Each of these subsystems will be discussed extensively in Chapter 4 from the architectural aspect of the component, and in Chapter 5 from a deployment point of view. 
\paragraph{}
It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show its integration into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption. This information about consumption is produced and communicated periodically. The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}. These commands request the sensor node to turn On or Off an appliance.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system. This was achieved by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world. This software tool is perceived as a heterogeneity handler, and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.

\section{Outline of the Thesis}
\paragraph{}
The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
\paragraph{Chapter 1}
It is an introductory chapter that gives a flavor of the project that is  studied in this thesis and also introduces the reader to the problem and the solution proposed in this context. This chapter also puts into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 
\paragraph{Chapter 2}
It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}. It also discusses related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ. Finally, its presents the commonalities ad differences between this thesis and the work previously done.
\paragraph{Chapter 3}
This part is dedicated to \gls{iot}, its importance and its impact on the world. Is also presents the actual state of the Internet and makes predictions about its future.
\paragraph{Chapter 4}
It presents the architecture of the test-bed. It also discusses the system from different perspectives, and how the problems stated in section 1.3 were solved from the architectural aspect.
\paragraph{Chapter 5}
This chapter mostly discusses the system deployment. It starts by presenting the hardware used in this project, then moves to the technology enablers. Afterwards, each component of the system is discussed extensively and moves along with the integration to form the test-bed.
\paragraph{Chapter 6}
It starts first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered is shown, analyzed and interpreted.
\paragraph{Chapter 7, 8}
These are concluding chapters of this study. They provide directions about future work, recommendations to the reader about what parts of the system could be improved and how these improvements could be achieved.

@@

The actual electrical grid is based on a technology that is at least one century old and its weaknesses are becoming more and more costly \cite{ref5}. Therefore, there is a need to change it to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy. As a matter of fact, research introduced us to a new technology: the \glsreset{sg}\gls{sg}. Although it is still a research project that has not reached the stage of maturity yet, nations across the globe started to deploy \gls{sg} projects. The \gls{sg} is relying on Information Technology to solve many of the problems existing in the traditional grid. Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.

@@

The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduces the losses of energy in the transmission.  This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer. Households can now produce and sell the surplus of electricity. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide electricity to consumers in a "free market" fashion which means that it will obey the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}

@@

The \glsreset{iot}\gls{iot} has been highly accepted by the players in the Internet as its adoption is starting smoothly. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make it part of it. Consequently, research has taken the direction of integrating the \gls{sg} within the \gls{iot} as most of the aspects of this last favor  the \gls{sg} \cite{ref8}.

@@

\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}. The smart grid's architecture based on \gls{iot} is divided into three layers:

@@

\item Perception layer: Devices autonomously perform their tasks such as sensing and executing orders.
\item Communication layer: The communication details are clarified here. It hides the communication issues and provide an easy for upper layers to send and receive information.
\item Application layer: Applications are built to exploit the data received from the underlying layers.

@@

Research discusses the benefits to replace Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10}. Still, Zigbee has the exclusive advantage of consuming the least energy of the communication technologies that exist nowadays which is something required in \gls{sg} and \gls{iot}, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices. 

@@

From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues. An \gls{iot} gateway system based on Zigbee and GPRS protocols helps to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network \cite{ref11}. Another solution to the heterogeneity problem is proposed with a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices. This protocol is platform-independent, low energy communication \cite{ref12}. Some researchers claim that the major source of heterogeneity the fact that there are different types of \gls{wsn} devices (Micaz, Mica2, Telosb...) that do not use the same standards and protocols. They propose to move the communication in \glspl{wsn} to an "all-IP" as this would remove most of the heterogeneity but as there are already deployed legacy \glspl{wsn}, the task is complicated to conduct. Fortunately, the researchers sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6} \cite{ref13}.\glsreset{sg}

\chapter{Internet Of Things}
\section{The Actual Internet}
\paragraph{}
The \glsreset{iot}\gls{iot} refers to the evolution of the actual Internet.The term \gls{iot} was used for the first time by Kevin Ashton in 1999 \cite{ref17}. The human beings will represent in the future a very small minority in the Internet and as it will be rules by autonomous machines. \gls{iot} will become the biggest machine ever created by human beings as the smart grid will be part of it.
\section{Introduction to \gls{iot}}
\paragraph{}
The \glsreset{iot}\gls{iot} is a network mostly ruled my machines, from the most powerful supercomputer and mainframes to the smallest devices such as sensor nodes. All will have one common point, being all connected to the Internet and able to communicate, produce and consume information. The \gls{iot} is changing the world of computing. New paradigms, architectures, frameworks and prototypes are built and refined everyday to support it \cite{ref15}. New technologies such as \glsreset{rfid}\gls{rfid}, \glsreset{m2m}\gls{m2m}, \glsreset{ipv6}\gls{ipv6}, \glsreset{6lowpan}\gls{6lowpan} and \gls{ims} have come to place to establish the infrastructure for \gls{iot}. The whole world is now being connected including humans and objects \cite{ref18}.
\paragraph{}
The \glsreset{iot}\gls{iot} brings numerous advantages but many issues will have to solved before this network explodes to connecting trillions. The first issue is the interconnection of all the devices into the \gls{iot}. What hardware will support such an ultra high traffic network with an ultra high performance and availability requirements. Will ubiquitous computing or pervasive computing have architectures that deal with such infrastructure?
\paragraph{}
To better understand \gls{iot},The following are the three steps in its flow of information:
\begin{enumerate}
\item The object is identified by making use of \gls{rfid}. In addition, the objects should be recognized by sensing the environment and this information should be converted to electronic format
\item  The information is transported to the information consumer.
\item The information is processed intelligently \cite{ref17}.
\end{enumerate}
